Associahedra for finite type cluster algebras \\ and minimal relations between $\b{g}$-vectors

### 2019-07-09 thirdPaper/draft/typeConeAssociahedra.tex
main results gentle

\subsection*{Cluster algebras of finite type} (\cref{subsec:typeConeCA})

We first consider the~$\b{g}$-vector fan~$\Fan(\B_\circ)$ of a skew-symmetrizable cluster algebra of finite type, with respect to any initial exchange matrix of finite type~$\B_\circ$.
A reformulation of~\cite[Thm 7.5]{BuanMarshReinekeReitenTodorov}
\medskip

\noindent \ref{prop:uniqueExchangePropertyCA}: The cluster fan~$\Fan(\B_\circ)$ has the unique exchange relation property.

@@

We consider the set~$\meshes$ of all pairs~$\{x,x'\}$ related by non-initial mesh mutations (\cref{def:meshMutation}), and we denote by~$\principalVariables$ the set of cluster variables.
Let~$\{x,x'\}\in\meshes$, and let~$X,X'$ be two clusters such that~$X\setminus\{x\}=X'\setminus\{x'\}$.
For any~$y\in\principalVariables$, define~$\alpha_{x,x',y}$ to be~$|b_{xy}|$ if~$y\in X\cap X'$ and 0 otherwise.
By the unique exchange relation property, the~$\alpha_{x,x',y}$'s do not depend on the choice of the clusters~$X,X'$.

@@

\subsection*{Gentle algebras}

\subsection*{Cluster categories}

\subsection*{Extriangulated categories}

@@

\begin{corollary}
\label{coro:simplicialTypeConeNKC}
For any brick and $2$-acyclic gentle quiver~$\quiver$, the type cone~$\typeCone \big( \gvectorFan[\quiver] \big)$ of the non-kissing fan~$\quiver$ is simplicial.
\end{corollary}

@@

\begin{corollary}
\label{coro:allPolytopalRealizationsNKC}
For any brick and $2$-acyclic gentle quiver~$\quiver$ and any~$\b{\ell} \in \R_{>0}^{\strings}$, the polytope
\[
R_\b{\ell}(\quiver) \eqdef \set{\b{z} \in \R^{\walks}}{\begin{array}{l} \b{z} \ge 0 \qquad\text{and}\qquad \b{z}_\omega = 0 \text{ for any improper walk } \omega \\ \b{z}_{\cc{\sigma}} + \b{z}_{\hh{\sigma}} - \b{z}_{\hc{\sigma}} - \b{z}_{\ch{\sigma}} = \b{\ell}_\sigma \text{ for all } \sigma \in \strings\end{array}}
\]
is a realization of the non-kissing fan~$\gvectorFan[\quiver]$.
Moreover, the polytopes~$R_\b{\ell}(\quiver)$ for~$\b{\ell} \in \R_{>0}^{\strings}$ describe all polytopal realizations of~$\gvectorFan[\quiver]$.
\end{corollary}

>>>>

@@

\noindent {\bf Cluster algebras of finite type} (\cref{subsec:typeConeCA})

We first consider the~$\b{g}$-vector fan~$\Fan(\B_\circ)$ of a skew-symmetrizable cluster algebra of finite type, with respect to any initial exchange matrix of finite type~$\B_\circ$.
A reformulation of~\cite[Thm 7.5]{BuanMarshReinekeReitenTodorov} gives the first step towards applying our type cone strategy.

\noindent \cref{prop:uniqueExchangePropertyCA}: \emph{The cluster fan~$\Fan(\B_\circ)$ has the unique exchange relation property.}

The second step follows from our main result in \cref{sec:clusterCategories}, which makes use of representation theory of finite dimensional algebras and additive categorification of cluster algebras.

\noindent \cref{coro:simplicialTypeConeCA}: \emph{The type cone~$\typeCone(\Fan(\B_\circ))$ is simplicial.}

We consider the set~$\meshes$ of all pairs~$\{x,x'\}$ related by non-initial mesh mutations (\cref{def:meshMutation}), and we denote by~$\principalVariables$ the set of cluster variables.
Let~$\{x,x'\}\in\meshes$, and let~$X,X'$ be two clusters such that~$X\setminus\{x\}=X'\setminus\{x'\}$.
For any~$y\in\principalVariables$, define~$\alpha_{x,x',y}$ to be~$|b_{xy}|$ if~$y\in X\cap X'$ and 0 otherwise.
By the unique exchange relation property, the~$\alpha_{x,x',y}$'s do not depend on the specific choice of the clusters~$X,X'$.

@@

\bigskip

\noindent {\bf Gentle algebras, non-kissing complex, generalized accordiohedra} (\cref{subsec:typeConeNKC})

The same strategy is applied to a class of finite dimensional algebras, called gentle algebras, whose representations are combinatorially well understood.
We identify two conditions, called brick and 2-acyclicity (\cref{def:brick2acyclic}), that enable us to make use of the type cone approach.
We note that the~$\tau$-tilting theory of gentle algebras is the algebraic counterpart of the combinatorics of accordions on surfaces~\cite{PaluPilaudPlamondon-nonkissing,BrustleDouvilleMousavandThomasYildirim,PaluPilaudPlamondon-surfaces}.

Let~$\quiver = (Q,I)$ be a gentle bound quiver, and let~$\Fan(\quiver)$ be the fan of its~$\b{g}$-vectors (\cref{subsubsec:nonkissingComplex}), also called its non-kissing fan.
In \cref{sect::Kbproj}, we make use of representation theory in order to prove:

\noindent \cref{prop:exchangeablePairsNKC} (ii): \emph{The non-kissing fan~$\Fan(\quiver)$ has the unique exchange relation property.}

\cref{subsubsec:simplicialTypeConeNKC} is dedicated to giving a purely combinatorial proof of:

\noindent \cref{coro:simplicialTypeConeNKC}: \emph{Assume that the gentle bound quiver~$\quiver$ is brick and 2-acyclic.
Then the type cone~$\typeCone(\Fan(\quiver))$ is simplicial.}

An alternative algebraic proof, using the main result of \cref{sec:extricats}, is given in \cref{sect::Kbproj}.

Write~$\strings$ for the set of all strings of~$\quiver$, and~$\walks$ for the set of all walks on~$\quiver$ (i.e. of all strings of the blossoming bound quiver of~$\quiver$ that join two blossom vertices).
If~$\sigma$ is a string of~$\quiver$, we make use of the evocative notations~$\cc{\sigma}, \hh{\sigma}, \hc{\sigma}, \ch{\sigma}$ for the walks obtained by adding, in the blossoming bound quiver, hooks or cohooks at each end of~$\sigma$.
We make use of our type cone approach to deduce:

\noindent \cref{coro:allPolytopalRealizationsNKC}: \emph{For any brick and $2$-acyclic gentle quiver~$\quiver$ and any~$\b{\ell} \in \R_{>0}^{\strings}$, the polytope
\[
R_\b{\ell}(\quiver) \eqdef \set{\b{z} \in \R^{\walks}}{\begin{array}{l} \b{z} \ge 0 \qquad\text{and}\qquad \b{z}_\omega = 0 \text{ for any improper walk } \omega \\ \b{z}_{\cc{\sigma}} + \b{z}_{\hh{\sigma}} - \b{z}_{\hc{\sigma}} - \b{z}_{\ch{\sigma}} = \b{\ell}_\sigma \text{ for all } \sigma \in \strings\end{array}}
\]
is a realization of the non-kissing fan~$\gvectorFan[\quiver]$.
Moreover, the polytopes~$R_\b{\ell}(\quiver)$ for~$\b{\ell} \in \R_{>0}^{\strings}$ describe all polytopal realizations of~$\gvectorFan[\quiver]$.}

\bigskip

\noindent {\bf Cluster categories}

\bigskip

\noindent {\bf Extriangulated categories}

@@

\begin{corollary}
\label{coro:simplicialTypeConeNKC}
For any brick and $2$-acyclic gentle quiver~$\quiver$, the type cone~$\typeCone \big( \gvectorFan[\quiver] \big)$ of the non-kissing fan~$\gvectorFan[\quiver]$ is simplicial.
\end{corollary}

@@

\begin{theorem}
\label{coro:allPolytopalRealizationsNKC}
For any brick and $2$-acyclic gentle quiver~$\quiver$ and any~$\b{\ell} \in \R_{>0}^{\strings}$, the polytope
\[
R_\b{\ell}(\quiver) \eqdef \set{\b{z} \in \R^{\walks}}{\begin{array}{l} \b{z} \ge 0 \qquad\text{and}\qquad \b{z}_\omega = 0 \text{ for any improper walk } \omega \\ \b{z}_{\cc{\sigma}} + \b{z}_{\hh{\sigma}} - \b{z}_{\hc{\sigma}} - \b{z}_{\ch{\sigma}} = \b{\ell}_\sigma \text{ for all } \sigma \in \strings\end{array}}
\]
is a realization of the non-kissing fan~$\gvectorFan[\quiver]$.
Moreover, the polytopes~$R_\b{\ell}(\quiver)$ for~$\b{\ell} \in \R_{>0}^{\strings}$ describe all polytopal realizations of~$\gvectorFan[\quiver]$.
\end{theorem}
